Cover letter for the original IJRR submission

### 2017-10-10 reviews_tro/reviewer_response.tex

In the following, we list all the questions and remarks from the Reviewers, answer to the questions, and detail the modifications that were brought to the paper.
At the end of this letter, we also join a diff file between the original submission and the present one, that should help the Reviewers see the modifications 
brought to the paper.

\section{Answers to Reviewer 1}
%~ \ndone reecriture p3 + enlever main video

\quot {As for novel contributions, this did not seem to be addressed/explained
until page 3 (in the last paragraph of Sec. I). Much of the methodology
has already been presented in earlier work, and the particular aspects
that set THIS paper apart seem (as claimed by the authors) to be: 
}

\as{Thanks for the comment. As mentioned in the summary, our planner is almost identical to the one in [20], but the paper comprises important additions that are the pseudo code for the algorithm (the algorithm is not given in [20]), as well as the introduction of a more efficient criterion
for checking static equilibrium.  The main novelty of the paper resides in the experimental setup and analysis, as well as the empirical proof that the planner works with real robot models that are kinematically more constrained than the virtual avatars from [20]. This is now explained in details in the introduction.}
\qt{(Section I.B) The present paper is
an extension of our ISRR conference paper [20]. As such, our
solution to address $\mathcal{P}_1$  and $\mathcal{P}_2$  is the same motion planner
as the one presented at ISRR (reformulated in Section IV
and V). However three important novelties have been added to
the planner: the pseudo-code of the algorithm (Section V-B2),
a novel criterion for static equilibrium (Section VI), and the
release of the source code of the planner (Section VII). \\
The other novelty of the paper is a rigorous experimental
validation of the approach on actual robot models (Sec-
tion VIII). To validate our contact plans we introduced a
complete framework for multi-contact motion synthesis. This
framework additionally comprises an interpolation method
to solve the problem $\mathcal{P}_3$, based on a reformulation of our
previous work [1]. Our solution to $\mathcal{P}_3$ allows us to verify that
the synthesized motions are physically consistent, using our
implementation of a state-of-the-art simulation algorithm [28].
These aspects of the framework are presented in details in the
paper, but are not novel per se. The novelty lies in the complete
validation of the contact planner with real robot models, for
which the planning is much harder with respect to the avatars
used in [20], because of more restrictive kinematic constraints}
\done

@@

\quot{It is a pretty intuitive/typical check for quantifying a
margin for friction conesque violations (using an approximation for
the friction cone, with x and y direction forces considered
independently). }

\as{Thanks for the comment. To our knowledge, what is common in the literature is to reduce the friction coefficient to obtain more conservative bounds on the admissible forces, and thus a ``safety margin'' to the real cone boundary. An issue with this formulation is that the resulting margin may be too small for small forces, and too large for large forces. With our formulation instead, the margin is constant (i.e. the value $b_0$). We updated the paper to explain this difference. We believe this formulation is novel, but would be happy to cite any relevant reference if we happen to be wrong. This formulation proves useful to us because the margin $b_0$ allows to choose the more robust candidate for static equilibrium. }\done
\qt{(Section VI.2) A simple approach to robustness consists in choosing a smaller friction coefficient, to constrain the forces to lie away from the boundaries of the real cone.  However, this would result in a small safety margin for forces of low magnitude, and an excessively large safety margin for large forces as the boundaries grow more and more apart. In comparison, our margin $b_0$ is constant, and provides a helpful mean to compare the robustness of different contact configurations.}

\quot{
Provides complete pseudocode.  This is great! However,it's
not fully clear how much of this is novel vs just re-presenting the
same algorithms previously publilshed.	If only code in App. B and C
are new, that should be clear. Conversely, if most of this presentation
was NOT outlined in such detail before, that should also be clear (to
the authors' credit). Either way, a brief clarification in the intro is
recommended.}

\as{We agree with the comment. Indeed, the pseudocode was not present in [20]. We included it in the core of the paper and clarified that it is novel at the end
of Section I.B.}
\qt{ our
solution to address $\mathcal{P}_1$  and $\mathcal{P}_2$  is the same motion planner
as the one presented at ISRR (reformulated in Section IV
and V). However three important novelties have been added to
the planner: the pseudo-code of the algorithm (Section V-B2),
a novel criterion for static equilibrium (Section VI), and the
release of the source code of the planner (Section VII).}
\done

@@

\quot{(3)	The contribution to provide a solution to P3 doesn't happen
until page 13 (Appendix C) and as with contribution (1), it needs a
better, intuitive summary (briefly) when first mentioned, as well as a
roadmap for the reader as to where it will be described. It is also
ambiguous in the paper whether to what extent (if any) this solution is
novel, versus a presentation of work from [1]. "Provide a solution"
seemed to imply novelty, in the context of this paragraph -- yet later,
the authors (near the end of page 2) state they use a "state-of-the-art
solution to P3", which implies the P3 solution is not a claim of
novelty.}

\as{The main contribution of the paper is the proposal and validation of an algorithm for P1 and P2. 
In the new version of the paper, we do not consider that our solution to P3
is a new contribution, although our formulation differs from [1] (our problem variables are the accelerations
of the center of mass, whereas in [1] the problem variables are the contact forces). We provide it here for completeness, because it corresponds to our implementation, and 
was required to validate our approach. 
We rephrased the text to clearly indicate that the algorithm is not a contribution. We made the choice not to give an intuition nor describe the solver in the core
of the paper because it is not required to understand the contributions to P1 and P2.}
\qt{(Section I.B) To validate our contact plans we introduced a complete framework for multi-contact motion synthesis. This framework additionally comprises an interpolation method to solve the problem $\mathcal{P}_3$,  based on a reformulation of our previous work [1]. Our solution to $\mathcal{P}_3$ allows us to verify that the synthesized motions are physically consistent, using our implementation of a state-of-the-art simulation algorithm [28]. 
These aspects of the framework are presented in details in the
paper, but are not novel \textit{per se}.}\done

@@

\as{Thanks for the comment. Regarding the absence of details on the dynamic simulator, first of all we apologize because [38] is not the appropriate reference.
We added a section VIII.A where we describe the simulation in more details.

Regarding the difference between our previous and current work, the postures computed in [20] were not validated in a simulator, nor the complete motions since 
we did not have a solution to P3 at the time. Then the main
difference between the avatars of [20] and the real robots we now address are the joint limitations and auto-collisions probabilities.
The robots are much more constrained in this regard, which makes the planning harder. We thus had to demonstrate that our algorithm is also working
for these models, while maintaining good performance.
We added this missing piece of information to the paper, at the end of Section I.B.}
\qt{(Section VIII.A) To generate continuous movements from our contact plans we used either the framework proposed in [1], or our own implementation of a $\mathcal{P}_3$ solver (Appendix C).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in [1]), or with our dynamic simulator, based on a state-of-the-art algorithm [25].
In the simulations we controlled the robot with a standard inverse-dynamics controller[31]. The code source of the simulator is available at \url{https://github.com/andreadelprete/pinocchio_inv_dyn/releases/tag/rbprm}.
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.}
\done
%~ \textcolor{red}{TODO ANDREA}.

\quot{p.14 - not a good description of the simulation software used [\dots]Ref [38] focuses on robustness of simulations, but presumably, the testing in the present work is deterministic.
 just using the same simulator? Not much description is given of how/whether contacts/slipping are modeled [\dots] Why isn't a 3rd party software used, or at least software with better documentation ? Better info is needed to understand what the simulation truly represents.}
\as{Thanks for the comment. In summary, we had to implement our
simulator because we were not satisfied with off-the-shelf simulators. Our
simulator implements the algorithm proposed by the paper ``Staggered projections for frictional
contact in multibody systems'', by Kauffman, Sueda, James and Pai, which
corresponds to a state-of-the-art solution to the problem.
We considered other simulators, namely Mujoco, Gazebo and
Open-HRP. We discarded Mujoco because it only provides a Windows 
implementation, while we work on Linux. In our tests with Gazebo, we were not able to
obtain coherent results, even for simple pendulum models. 
Finally, Open-HRP is a great tool, but it only provides an interface for 
position-controlled models, and not torque-controlled. For these reasons, 
and other historical ones, we implemented our own simulator. 
We understand that because we use an
in-house implementation, it is harder to trust the results we obtain.
However, we implemented a state of the art LCP algorithm, used
by many in the community. In this new submission, we also provide the
source of the simulator.}

@@

\section{Answers to Reviewer 6}
\quot{ My primary concern, however,
is that the elements described in this paper have been described in
previous published work, which makes the particular contributions of
this work unclear. To be specific, the reachability-guided planner
appears identical to the planner in [20] (as stated by the authors in
the end of Section I.B.), and the optimal control formulation is from [1]
(again as stated in I.B.).}
\as{Thanks for the comment. Indeed, the Reviewers agreed that the contribution with respect to previous work was not clearly stated. Here we reproduce the answer we made to Reviewer I: As mentioned in the summary, our planner is almost identical to the one in [20], but the paper comprises important additions that are the pseudo code for the algorithm (the algorithm is not given in [20]), as well as the introduction of a more efficient criterion
for checking static equilibrium.  The main novelty of the paper resides in the experimental setup and analysis, as well as the empirical proof that the planner works with real robot models that are kinematically more constrained than the virtual avatars from [20]. This is now explained in details in the introduction.}
\qt{(Section I.B, repeated from page 5) The present paper is
an extension of our ISRR conference paper [20]. As such, our
solution to address $\mathcal{P}_1$  and $\mathcal{P}_2$  is the same motion planner
as the one presented at ISRR (reformulated in Section IV
and V). However three important novelties have been added to
the planner: the pseudo-code of the algorithm (Section V-B2),
a novel criterion for static equilibrium (Section VI), and the
release of the source code of the planner (Section VII). \\
The other novelty of the paper is a rigorous experimental
validation of the approach on actual robot models (Sec-
tion VIII). To validate our contact plans we introduced a
complete framework for multi-contact motion synthesis. This
framework additionally comprises an interpolation method
to solve the problem $\mathcal{P}_3$, based on a reformulation of our
previous work [1]. Our solution to $\mathcal{P}_3$ allows us to verify that
the synthesized motions are physically consistent, using our
implementation of a state-of-the-art simulation algorithm [28].
These aspects of the framework are presented in details in the
paper, but are not novel per se. The novelty lies in the complete
validation of the contact planner with real robot models, for
which the planning is much harder with respect to the avatars
used in [20], because of more restrictive kinematic constraints.}\done

### 2017-10-10 body/results.tex
rewrote I.B more like in the letter
link to simulator source code
final version of main document .?

\subsection{Experimental validation of the contact plans} \label{sec:valid}
%~ \adnote{I added the description of the simulations here, but I wonder whether this part should not be moved to the beginning of the Results section.}
To generate continuous movements from our contact plans we used either the framework proposed in \cite{Carpentier2016}, or our own implementation of a $\mathcal{P}_3$ solver (Appendix~\ref{app:optim}).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in~\cite{Carpentier2016}), or with our dynamic simulator, based on a state-of-the-art algorithm~\cite{Kaufman2008}.
In the simulations we controlled the robot with a standard inverse-dynamics controller~\cite{DelPrete2015b}.
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.

@@

% \subsubsection*{Truck egress (Figure~\ref{res_truck_pres} and Figure~\ref{res_truck_bd}) -- Humanoid and insectoid robots.}
%\subsubsection*{Truck egress -- Humanoid and insectoid robots (Figure~\ref{res_truck_bd}).}
\subsubsection{HRP-2 -- Staircase with high steps (Figure~\ref{fig:stair_robust})}
\newcommand{\widthValue}{0.13\linewidth}

\begin{figure*}[h!t!]
  \begin{center}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing1.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing2.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing3.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing4.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing5.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing6.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing7.jpeg}
  \caption{
           HRP-2 in the stair climbing scenario. }
		   \label{fig:stair_robust}
  \label{fig:moviepicture2}
  \end{center}
  \vspace{-0.5cm}
\end{figure*}

@@

@@

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus allowing to consider \gls{interactive} applications and online autonomous planning of the robot motion.

>>>>

\subsection{Experimental validation of the contact plans} \label{sec:valid}
%~ \adnote{I added the description of the simulations here, but I wonder whether this part should not be moved to the beginning of the Results section.}
To generate continuous movements from our contact plans we used either the framework proposed in \cite{Carpentier2016}, or our own implementation of a $\mathcal{P}_3$ solver (Appendix~\ref{app:optim}).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in~\cite{Carpentier2016}), or with our dynamic simulator, based on a state-of-the-art algorithm~\cite{Kaufman2008}.
In the simulations we controlled the robot with a standard inverse-dynamics controller~\cite{DelPrete2015b}. The code source of the simulator is available at \url{https://github.com/andreadelprete/pinocchio_inv_dyn/releases/tag/rbprm}.
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.

@@

% \subsubsection*{Truck egress (Figure~\ref{res_truck_pres} and Figure~\ref{res_truck_bd}) -- Humanoid and insectoid robots.}
%\subsubsection*{Truck egress -- Humanoid and insectoid robots (Figure~\ref{res_truck_bd}).}

@@

\subsubsection{HRP-2 -- Staircase with high steps (Figure~\ref{fig:stair_robust})}
\newcommand{\widthValue}{0.13\linewidth}

\begin{figure*}[h!t!]
  \begin{center}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing1.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing2.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing3.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing4.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing5.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing6.jpeg}
  \includegraphics[trim={7.0cm 0.0cm 20.0cm 0.0cm}, clip, width=\widthValue]
    {./video/stairclimbing7.jpeg}
  \caption{
           HRP-2 in the stair climbing scenario. }
		   \label{fig:stair_robust}
  \end{center}
  \vspace{-0.5cm}
\end{figure*}

@@

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus enabling \gls{interactive} applications and online autonomous planning of the robot motion.
